---
title: "Activity Prediction"
author: "Jeremy Sellars"
date: "September 27, 2015"
output: html_document
---

# Executive summary

This project fits a random forest to activity tracker data from http://groupware.les.inf.puc-rio.br/har and uses it to predict the type of activity being performed based on never-before-seen data.  An analysis of the expected out-of-sample error rate is included.

## Fitting the model

Some columns were blank or NA (Not available) and these were removed.  Also, the username and sample numbers were removed.  The other columns were considered in the analysis.

```{r cache=T, echo=F}
setwd("C:/code/jeremy.sellars/coursera/predmachlearn/project")
suppressMessages(suppressWarnings({
  library(knitr)
  library(caret)
  library(randomForest)
  }))
```

```{r cache=T}
training <- read.csv("pml-training.csv")
training <- training[, !(names(training) %in% c("X","user_name","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","new_window","num_window","kurtosis_roll_belt","kurtosis_picth_belt","kurtosis_yaw_belt","skewness_roll_belt","skewness_roll_belt.1","skewness_yaw_belt","max_roll_belt","max_picth_belt","max_yaw_belt","min_roll_belt","min_pitch_belt","min_yaw_belt","amplitude_roll_belt","amplitude_pitch_belt","amplitude_yaw_belt","var_total_accel_belt","avg_roll_belt","stddev_roll_belt","var_roll_belt","avg_pitch_belt","stddev_pitch_belt","var_pitch_belt","avg_yaw_belt","stddev_yaw_belt","var_yaw_belt","var_accel_arm","avg_roll_arm","stddev_roll_arm","var_roll_arm","avg_pitch_arm","stddev_pitch_arm","var_pitch_arm","avg_yaw_arm","stddev_yaw_arm","var_yaw_arm","kurtosis_roll_arm","kurtosis_picth_arm","kurtosis_yaw_arm","skewness_roll_arm","skewness_pitch_arm","skewness_yaw_arm","max_roll_arm","max_picth_arm","max_yaw_arm","min_roll_arm","min_pitch_arm","min_yaw_arm","amplitude_roll_arm","amplitude_pitch_arm","amplitude_yaw_arm","kurtosis_roll_dumbbell","kurtosis_picth_dumbbell","kurtosis_yaw_dumbbell","skewness_roll_dumbbell","skewness_pitch_dumbbell","skewness_yaw_dumbbell","max_roll_dumbbell","max_picth_dumbbell","max_yaw_dumbbell","min_roll_dumbbell","min_pitch_dumbbell","min_yaw_dumbbell","amplitude_roll_dumbbell","amplitude_pitch_dumbbell","amplitude_yaw_dumbbell","var_accel_dumbbell","avg_roll_dumbbell","stddev_roll_dumbbell","var_roll_dumbbell","avg_pitch_dumbbell","stddev_pitch_dumbbell","var_pitch_dumbbell","avg_yaw_dumbbell","stddev_yaw_dumbbell","var_yaw_dumbbell","kurtosis_roll_forearm","kurtosis_picth_forearm","kurtosis_yaw_forearm","skewness_roll_forearm","skewness_pitch_forearm","skewness_yaw_forearm","max_roll_forearm","max_picth_forearm","max_yaw_forearm","min_roll_forearm","min_pitch_forearm","min_yaw_forearm","amplitude_roll_forearm","amplitude_pitch_forearm","amplitude_yaw_forearm","var_accel_forearm","avg_roll_forearm","stddev_roll_forearm","var_roll_forearm","avg_pitch_forearm","stddev_pitch_forearm","var_pitch_forearm","avg_yaw_forearm","stddev_yaw_forearm","var_yaw_forearm"
))]

set.seed(654798)  # Reproducible data
inTrain <- createDataPartition(y=training$classe, p=.3, list=F)  # Use small 30% training set for sake of computation time on my slow computer
train <- training[inTrain,]
test <- training[-inTrain,]

modFit <- train(classe~ ., data=train,method="rf",prox=T, ntree=10)

pred <- predict(modFit,test)
test$predRight <- pred==test$classe
```

### Cross Validation and Out of Sample Rate
`r kable(table(pred,test$classe))`

```{r echo=T}
t <- table(pred,test$classe)
total_obs = sum(t)
for(n in 1:nrow(t)) t[n,n] <- 0  # Set diagonals to 0 to find mislabel count
mislabel_rate = sum(t) / total_obs
```

The out-of-sample error rate is `r round(mislabel_rate * 100, 2)`%

```{r eval=F, echo=F}
# This section creates submission files and is not a part of the document.
submission_test <- read.csv("pml-testing.csv")
answers = predict(modFit,submission_test)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)
```
